---
title: "Lab1Intro"
output: 
    rmarkdown::html_vignette:
        toc: true
vignette: >
  %\VignetteIndexEntry{Lab1Intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab1Intro)
```

# Links

The github repo can be found [here](https://github.com/jeffniv/Lab1Intro). The documentation for making a packing can be found [here](http://r-pkgs.had.co.nz/). 

# Definitions

The *sample mean* can be computed from the $n$ measurements on each of the $p$ variables, so that, in general, there will be $p$ sample means:
$$\bar{x}_k = \frac{1}{n} \sum^n_{j=1} x_{jk}, \ \ k=1,2,\ldots,p.$$

The *sample covariance* measures the association between the ith and kth variables:
$$s_{ik} = \frac{1}{n} \sum^n_{j=1} (x_{ji} - \bar{x}_i)(x_{jk} - \bar{x}_k), \ \ i=1,2,\ldots,p, \ k=1,2,\ldots,p.$$ 

The *sample correlation* coefficient for the ith and kth variables is defined as:
$$r_{ik} = \frac{s_{ik}}{\sqrt{s_{ii}}\sqrt{s_{kk}}} = \frac{\sum^n_{j=1} (x_{ji} - \bar{x}_i)(x_{jk} - \bar{x}_k)}{\sqrt{\sum^n_{j=1} (x_{ji} - \bar{x}_i)^2} \sqrt{\sum^n_{j=1} (x_{jk} - \bar{x}_k)^2}}$$.

Let bolded items represent vectors and matrices. $\mathbf{1}$ represent an $n\times 1$ vector with entries $1$. The *sample mean vector* is defined as:
$$\bar{\mathbf{x}} = \frac{1}{n}\mathbf{X}'\mathbf{1}$$

The *sample covariance matrix* is:
$$\mathbf{S} = \frac{1}{n-1} \mathbf{X}' \left(\mathbf{I} - \frac{1}{n}\mathbf{1}\mathbf{1}'\right)\mathbf{X} $$.

Once $\mathbf{S}$ is computed then the *sample standard deviation matrix* is defined as
$$\mathbf{D}^{1/2}_{(p \times p)} = \left[ \begin{array}{cccc} \sqrt{s_{11}} & 0 & \cdots & 0 \\
0 & \sqrt{s_{22}} & \cdots & 0 \\
\vdots & \vdots & \ddots & \vdots \\
0 & 0 & \cdots & \sqrt{s_{pp}} \end{array} \right]$$

Then
$$\mathbf{D}^{-1/2}_{(p \times p)} = \left[ \begin{array}{cccc} \frac{1}{\sqrt{s_{11}}} & 0 & \cdots & 0 \\
0 & \frac{1}{\sqrt{s_{22}}} & \cdots & 0 \\
\vdots & \vdots & \ddots & \vdots \\
0 & 0 & \cdots & \frac{1}{\sqrt{s_{pp}}} \end{array} \right]$$

Since
$$\mathbf{S} = \left[ \begin{array}{cccc} s_{11} & s_{12} & \cdots & s_{1p} \\
\vdots & \vdots & \ddots & \vdots \\
s_{1p} & s_{2p} & \cdots & s_{pp} \end{array} \right]$$

and
$$\mathbf{R} = \left[ \begin{array}{cccc} \frac{s_{11}}{\sqrt{s_{11}}\sqrt{s_{11}}} & \frac{s_{12}}{\sqrt{s_{11}}\sqrt{s_{22}}} & \cdots & \frac{s_{1p}}{\sqrt{s_{11}}\sqrt{s_{pp}}} \\
\vdots & \vdots & \ddots & \vdots \\
\frac{s_{1p}}{\sqrt{s_{11}}\sqrt{s_{pp}}} & \frac{s_{2p}}{\sqrt{s_{22}}\sqrt{s_{pp}}} & \cdots & \frac{s_{pp}}{\sqrt{s_{pp}}\sqrt{s_{pp}}} \end{array} \right] = \left[ \begin{array}{cccc} 1 & r_{12} & \cdots & r_{1p} \\
\vdots & \vdots & \ddots & \vdots \\
r_{1p} & r_{2p} & \cdots & 1 \end{array} \right]$$

We have
$$\mathbf{R} = \mathbf{D}^{-1/2}\mathbf{S}\mathbf{D}^{-1/2}$$.

# Examples

The package comes with paper quality data called `paper`.

```{r paper}
data(paper)
head(paper)
```

The function `mymeanvector()` calculates the mean vector of a data matrix.
```{r mean}
mymeanvector(paper)
```

The function `mycovariance()` calculates the biased covariance matrix of a data matrix.
```{r cov}
mycovariance(paper)
```

The function `mycorrelation()` calculates the correlation matrix of a data matrix.
```{r cor}
mycorrelation(paper)
```

# Assessment for this course

1. Clicker quizzes (10%)
2. 4 assignments (20%)
3. Laboratories (10%)
4. 2 mid-term exams (20%)
5. 3 projects (10%)
6. 1 final exam (30%)
